---
title: "Day 2: Data Types"
date: "15 August 2023"
format: html
editor: visual
execute: 
  warning: false
---

```{r}
library(tidyverse)
library(here)
```

## Making & Indexing Data in R

#### Vectors

```{r}
dogs <- c("teddy", "khora", "waffle", "banjo", "tallie")

# check the class of dogs
typeof(dogs)
class(dogs)
# sometimes these two will have different results; most of the time will use class()
```

```{r}
my_breakfast <- c("waffles", 2, TRUE, "coffee")
class(my_breakfast)
# different data types, so all coerced to character since it is the most flexible type
```

```{r}
weights <- c(50, 60, 25, 35, 55)
class(weights)
typeof(weights)
# notice that these have different outputs. they are stored as doubles- higher 64-bit
```

```{r}
weights_int <- c(50L, 60L, 25L, 35L, 55L)
typeof(weights_int)
class(weights_int)
# adding L after the numbers makes them integers
```

use "is.\_\_" functions to ask what class something is

```{r}
is.numeric(weights_int)
```

use "as.\_\_" functions to recast data classes

```{r}
zipcodes <- c(93105, 93514, 93101, 93401)
class(zipcodes)

zipcodes_chr <- as.character(c(93105, 93514, 93101, 93401))
class(zipcodes_chr)
```

#### Creating a vector with named elements

```{r}
dog_food <- c(teddy = "purina", khora = "alpo", waffle = "fancy feast", banjo = "blue diamond", tallie = "farmer's dogs")
class(dog_food)
# class will only refer to the values, even though this is a string- doesn't look that it is named
```

##### Access 2nd element in dog_food

```{r}
dog_food[2]
dog_food["khora"]
# R uses a 1 index- first element is given positional value of 1
# Python and JavaScript use 0-indexing
```

##### Access range of values in a vector

```{r}
trees <- c("ponderosa", "lodegpole", "jeffery", "sugar", "white", "bristlecone")
trees[3]
trees[3:5]
```

## For loops

-   iterates over different elements in a data structure; does something to each of those values and returns something new

```{r}
# creating an indexing helper
i <- 4
trees[i]
```

```{r}
i <- seq(1:3)
trees[i]
# gives us values for positions 1 through 3
```

##### Re-asign values based on position

```{r}
trees[3] <- "oak"
trees
```

## Matrix indexing

```{r}
fish_size <- matrix(c(0.8, 1.2, 1.4, 0.9), ncol = 2, nrow = 2, byrow = FALSE)
fish_size
class(fish_size)
str(fish_size)

fish_size[2,2]
fish_size[,2] # having the blank space there means keep all the rows
```

## Lists

```{r}
urchins <- list(color = "blue", values = c(1, 2, 3), meal = c("lunch", "dinner", "dessert"), cost = 5L)
urchins
# see we can have different data types and different data structures within a list
# don't have to give each element names, but you can if you want to and can be useful to refer to them later
```

```{r}
urchins_list <- urchins[2]
urchins_list
class(urchins_list)
# returns data but retains structure
urchins_vector <- urchins[[2]]
urchins_vector
class(urchins_vector)
# returns a vector
```

```{r}
urchins$values
```

```{r}
sample1 <- c(4,7,2,5,7,2,4)
sample2 <- c(3,2,7,8)

ttest_ex <- t.test(sample1, sample2)
ttest_ex
```

The output of a t-test is in a list even though it looks kinda funny in the output. So, we can pull values out of the list by their names.

```{r}
ttest_ex$p.value
ttest_ex$conf.int
```

Mean size for trees in Site A is similar to Site B (p = `r ttest_ex$p.value`).

## Data Frames

```{r}
fruit <- data.frame(type = c("apple", "banana", "peach"),
                    mass = c(130, 195, 150))
tibble::as_tibble(fruit)

fruit[3,2]
```

#### Using the pluck() function from purrr

-   helps with indexing from data structures

```{r}
purrr::pluck(fruit, "type", 2)
# (data, column, 2nd element)
```

```{r}
tibble::as_tibble(starwars)

starwars$starships[[1]]
# gets values
starwars$starships[1]
# gets list

starwars$starships[[1]][2]
purrr::pluck(starwars, "starships", 1, 2)
# getting 2nd value 
# these two lines both do the same thing
```

